Extreme Digitisation For Ground-Based Cosmic Microwave Background Experiments

### 2019-05-29 mnras_template.tex
Add files via upload

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[apj]{emulateapj}
%\documentclass[preprint2]{aastex61}
%\documentclass[12pt,preprint]{aastex}
\graphicspath{{figures/}}
\DeclareGraphicsExtensions{.jpg,.pdf,.png,.eps,.ps}

\usepackage[table,usenames,dvipsnames]{xcolor}
\usepackage{amsmath}
\usepackage{subfigure}
\usepackage[backref,breaklinks,colorlinks,citecolor=blue]{hyperref}
\usepackage{natbib}
%\usepackage{natbib}
\bibliographystyle{fapj}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{soul}

%\newcommand{\jcap}{JCAP}

@@

\newcommand{\tbd}[1]{\textcolor{Red}{{\bf TBD}: #1}}
\newcommand{\gab}[1]{\textcolor{Orchid}{[{\bf GS}: #1]}}
\newcommand{\changed}[1]{\textcolor{Red}{#1}}
\newcommand{\removed}[1]{\textcolor{Red}{}}
\include{number_list}

%

% ref to section \S\ref{sec:label}

%\submitjournal{ApJ}
\def\Melbourne{1}
\def\uci{2}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

\title{Extreme Digitisation For Ground-Based Cosmic Microwave Background Experiments}
\author{L.~Balkenhol\altaffilmark{\Melbourne} and C.~L.~Reichardt\altaffilmark{\Melbourne}}
\altaffiltext{\Melbourne}{School of Physics, University of Melbourne, Parkville, VIC 3010, Australia}
\email{lbalkenhol@student.unimelb.edu.au}

\begin{abstract}

The large size of the time ordered data of cosmic microwave background experiments presents challenges for mission planning and data analysis. These issues are particularly significant for Antarctica- and space-based experiments, which depend on satellite links to transmit data. We explore the viability of reducing the time ordered data to few bit numbers to address these challenges. Unlike lossless compression, few bit digitisation introduces additional noise into the data. We present a set of one, two, and three bit digitisation schemes and measure the increase in noise in the cosmic microwave background temperature and polarisation power spectra. The digitisation noise is independent of angular scale and is well-described as a constant percentage of the original detector noise. Three bit digitisation increases the map noise level by $< 2\%$, while reducing the data volume by a factor of \changed{ten relative to 32-bit floats. Lossless compression of the digitised data does not lead to additional data volume reduction.} Extreme digitisation is a promising strategy for upcoming experiments.

\end{abstract}

%C: maybe sub techniques:misc for methods: data analysis ? unfortunately no data compression tag
\keywords{cosmic background radiation --- polarization --- techniques: miscellaneous}
\section{Introduction}
\label{sec:intro}

@@

As seen in Figure \ref{fig:psrecover} (appendix) all spectra of extremely digitised data recover the input to a satisfying degree. Differences between few bit and the input power spectra are due to observing a single CMB realisation and residual boundary effects. \changed{Non-linearity introduced by the digitisation process disappears in the final map due to the large number of TOD points that contribute to each map pixel. Thus the digitisation effects are not an issue for power spectrum estimation.}

@@

\changed{Current-generation ground-based CMB experiments employ a combination of lossly, e.g. downsampling, and lossless compression, e.g. FLAC, bzip, LZMA, to manage their transmission bottlenecks. We test whether the data volume of timestreams that have undergone few-bit digitisation can be reduced further by applying the most popular lossless compression algorithm: FLAC. }

\changed{We obtain 2 minutes of SPT-3G TOD and apply the digitisation schemes in equations \ref{eq:1bit}-\ref{eq:3bit} before compressing the data further using FLAC. We compare the file size of FLAC compressed few-bit timestreams to the typical data reduction rate achieved for 24-bit SPT-3G data.}

\changed{We find that using FLAC in conjunction with extreme digitisation does not yield additional data volume reduction. For 3-bit digitisation, both the FLAC and raw timestream have $\sim$3 bits per data point. In fact, due to the overhead required by FLAC compressed files perform marginally worse than uncompressed few-bit streams. This is unsurprising, given that we have already maximally compressed the data using extreme digitisation. For comparison, FLAC reduces the full 24-bit  SPT-3G data to $6-8$ bits of entropy per number. For this real-world case, three-bit digitisation reduces the data volume by a factor of $2-3$.}

\section{Conclusion}
\label{sec:conclusions}

@@

We present a set of one, two, and three bit digitisation schemes. For white and $1/f$ detector noise alike, we find that optimal three bit digitisation adds as little as $<2\%$ to the map noise level. This is true for temperature and polarisation observations. No change in the results is observed for maps with different numbers of hits per pixel and no angular scale dependence was observed in the added noise. \changed{Combining few-bit digitisation with FLAC compression does not yield additional data volume reduction.}

@@

\acknowledgments % thank patrick for discussion/starting point? Andrew?

% Melbourne CMB group, funding, NERSC.

\changed{We are indebted to Nathan Whitehorn for the details about SPT data compression provided by him.}
We are grateful for insightful discussions about the prospects of extreme digitisation for CMB experiments with Andrew Melatos and Patrick Clearwater.
We thank the referee as well as Nikhel Gupta, Srinivasan Raghunathan, Federico Bianchini, Andrew Melatos, and Patrick Clearwater for valuable feedback on the manuscript.
We acknowledge support from an Australian Research Council Future Fellowship (FT150100074), and also from the University of Melbourne. 
This research used resources of the National Energy Research Scientific Computing Center, which is supported by the Office of Science of the U.S. Department of Energy under Contract No. DE-AC02-05CH11231. 
We acknowledge the use of the Legacy Archive for Microwave Background Data Analysis (LAMBDA). Support for LAMBDA is provided by the NASA Office of Space Science.

% all code stuff

This research made use of the NumPy \citep{numpy}, SciPy \citep{scipy}, Matplotlib \citep{matplotlib}, and Astropy \citep{astropy} packages.

\newpage

% APPENDIX
\appendix

\begin{figure*}[h!t]\centering
\vspace{-1.5cm}
\includegraphics[width=0.8\textwidth,clip]{psrecovery.pdf}
  \caption[Current ]{
  Input power spectra used to generate the CMB template map plus detector noise level (black line) and digitisation noise (grey line); power spectra using three- (violet marker, right), two- (orange marker, middle), and one-bit (green marker, left) TOD. Power spectra are rescaled to $D_\ell = \ell (\ell+1) C_\ell /2\pi$ and bined to $\Delta \ell = 100$. Main panels show the TT (top), EE (middle), and BB (bottom) power spectra; accompanying panels show the difference between the input plus noise curves and the power spectra recovered from few-bit TOD. Spectra constructed from few-bit TOD recover the input well. This plot shows results for a simulation with $1,024,000$ hits per map pixel using Gaussian white detector noise.}
\label{fig:psrecover}

@@

% result tables
\subsection{Digitisation Noise Levels}
\label{subsec:appendixnoisetables}

\def\arraystretch{1.3}
\begin{table*}[tbh]
\begin{center}
\caption{\label{tab:extranoisewhite} Fractional Noise Increase Due To Digitisation - Gaussian White Detector Noise}
\small
\begin{tabular}{c c c c c}
Hits Per Pixel & Channel & 1 Bit & 2 Bit & 3 Bit \\
\hline
\hline
\multirow{3}{*}{800}  & TT  & $ 0.252 \pm 0.0091 $  & $ 0.0639 \pm 0.0034 $  & $ 0.0173 \pm 0.0021 $ \\
& EE  & $ 0.253 \pm 0.012 $  & $ 0.0640 \pm 0.0047 $  & $ 0.0175 \pm 0.0041 $ \\
& BB  & $ 0.254 \pm 0.011 $  & $ 0.0652 \pm 0.0054 $  & $ 0.0180 \pm 0.0039 $ \\
\hline
\multirow{3}{*}{8,000}  & TT  & $ 0.2521 \pm 0.0076 $  & $ 0.0641 \pm 0.0034 $  & $ 0.0174 \pm 0.0017 $ \\
& EE  & $ 0.2519 \pm 0.0074 $  & $ 0.0636 \pm 0.0040 $  & $ 0.0173 \pm 0.0022 $ \\
& BB  & $ 0.252 \pm 0.018 $  & $ 0.0634 \pm 0.0089 $  & $ 0.0176 \pm 0.0036 $ \\
\hline
\multirow{3}{*}{80,000}  & TT  & $ 0.2526 \pm 0.0080 $  & $ 0.0645 \pm 0.0040 $  & $ 0.0181 \pm 0.0016 $ \\
& EE  & $ 0.2506 \pm 0.0093 $  & $ 0.0641 \pm 0.0042 $  & $ 0.0177 \pm 0.0019 $ \\
& BB  & $ 0.250 \pm 0.012 $  & $ 0.0636 \pm 0.0065 $  & $ 0.0179 \pm 0.0032 $ \\
\hline
\multirow{3}{*}{1,024,000}  & TT  & $ 0.2501 \pm 0.0090 $  & $ 0.0633 \pm 0.0031 $  & $ 0.0173 \pm 0.0016 $ \\
& EE  & $ 0.2518 \pm 0.0080 $  & $ 0.0646 \pm 0.0033 $  & $ 0.0179 \pm 0.0018 $ \\
& BB  & $ 0.2520 \pm 0.0082 $  & $ 0.0639 \pm 0.0037 $  & $ 0.0173 \pm 0.0021 $ \\
\hline
\multirow{3}{*}{10,240,000}  & TT  & $ 0.2480 \pm 0.0078 $  & $ 0.0630 \pm 0.0038 $  & $ 0.0175 \pm 0.0018 $ \\
& EE  & $ 0.2526 \pm 0.0077 $  & $ 0.0647 \pm 0.0033 $  & $ 0.0183 \pm 0.0018 $ \\
& BB  & $ 0.252 \pm 0.010 $  & $ 0.0642 \pm 0.0042 $  & $ 0.0177 \pm 0.0020 $ \\
\hline
\multirow{3}{*}{102,400,000}  & TT  & $ 0.2465 \pm 0.0049 $  & $ 0.0629 \pm 0.0019 $  & $ 0.0170 \pm 0.0010 $ \\
& EE  & $ 0.2511 \pm 0.0078 $  & $ 0.0644 \pm 0.0035 $  & $ 0.0175 \pm 0.0018 $ \\
& BB  & $ 0.2523 \pm 0.0086 $  & $ 0.0640 \pm 0.0031 $  & $ 0.0179 \pm 0.0017 $ \\
\hline
\multirow{3}{*}{Average}  & TT  & $ 0.2502 \pm 0.0017 $  & $ 0.0636 \pm 0.0017 $  & $ 0.0175 \pm 0.0017 $ \\
 & EE  & $ 0.2518 \pm 0.0023 $  & $ 0.0643 \pm 0.0023 $  & $ 0.0177 \pm 0.0023 $ \\
 & BB  & $ 0.2519 \pm 0.0027 $  & $ 0.0640 \pm 0.0027 $  & $ 0.0177 \pm 0.0027 $ \\
\end{tabular}
\tablecomments{ 
Fractional addition to the map noise level, $\Delta \sigma / \sigma$, due to one-, two-, and three-bit digitisation for white detector noise with $1\sigma$ confidence intervals. Note that there is no appreciable variation between temperature and polarisation observations, or between different depths of observation. It is striking that three-bit digitisation only leads to a percent-level increase in the map noise level.}
 \normalsize
\end{center}
\end{table*}

\def\arraystretch{1.3}
\begin{table*}[tbh]
\begin{center}
\caption{\label{tab:extranoiseoof} Fractional Noise Increase Due To Digitisation - $1/f$ Detector Noise}
\small
\begin{tabular}{c c c c c}
Hits Per Pixel & Channel & 1 Bit & 2 Bit & 3 Bit \\
\hline
\hline
\multirow{3}{*}{8,000}  & TT  & $ 0.231 \pm 0.017 $  & $ 0.0584 \pm 0.0052 $  & $ 0.0159 \pm 0.0022 $ \\
& EE  & $ 0.250 \pm 0.012 $  & $ 0.0641 \pm 0.0043 $  & $ 0.0177 \pm 0.0019 $ \\
& BB  & $ 0.245 \pm 0.021 $  & $ 0.0623 \pm 0.0067 $  & $ 0.0169 \pm 0.0032 $ \\
\hline
\multirow{3}{*}{102,400}  & TT  & $ 0.235 \pm 0.014 $  & $ 0.0597 \pm 0.0042 $  & $ 0.0166 \pm 0.0019 $ \\
& EE  & $ 0.2503 \pm 0.0091 $  & $ 0.0636 \pm 0.0042 $  & $ 0.0174 \pm 0.0016 $ \\
& BB  & $ 0.249 \pm 0.024 $  & $ 0.0641 \pm 0.0069 $  & $ 0.0178 \pm 0.0025 $ \\
\hline
\multirow{3}{*}{1,024,000}  & TT  & $ 0.238 \pm 0.013 $  & $ 0.0601 \pm 0.0046 $  & $ 0.0168 \pm 0.0019 $ \\
& EE  & $ 0.2502 \pm 0.0078 $  & $ 0.0630 \pm 0.0039 $  & $ 0.0173 \pm 0.0019 $ \\
& BB  & $ 0.2499 \pm 0.0095 $  & $ 0.0636 \pm 0.0049 $  & $ 0.0177 \pm 0.0019 $ \\
\hline
\multirow{3}{*}{Average}  & TT  & $ 0.2346 \pm 0.0020 $  & $ 0.0594 \pm 0.0020 $  & $ 0.0164 \pm 0.0020 $ \\
 & EE  & $ 0.2503 \pm 0.0018 $  & $ 0.0636 \pm 0.0018 $  & $ 0.0175 \pm 0.0018 $ \\
 & BB  & $ 0.2480 \pm 0.0025 $  & $ 0.0633 \pm 0.0025 $  & $ 0.0175 \pm 0.0025 $ \\
\end{tabular}
\tablecomments{ 
Fractional addition to the map noise level, $\Delta \sigma / \sigma$, due to one-, two-, and three-bit digitisation for $1/f$ detector noise with $1\sigma$ confidence intervals. As in the white noise case, we observe no significant trend between temperature and polarisation or for different depths of observation. We note that, even assuming a more realistic noise profile, three-bit digitisation only leads to a percent-level increase in the map noise level.}
 \normalsize
\end{center}
\end{table*}

@@

\begin{equation}
\begin{aligned}
\langle D(\mu + \xi_1) D(\mu + \xi_2) \rangle &= \left\{ \sum_{i=0}^N  \frac{y_i}{2} \left[ erf \left( \frac{x_{i+1} - \mu}{\sqrt{2}\sigma} \right) - erf \left( \frac{x_{i} - \mu}{\sqrt{2}\sigma} \right) \right] \right\} \\
&\times \left\{  \sum_{j=0}^N \frac{y_j}{2} \left[ erf \left( \frac{x_{j+1} - \mu}{\sqrt{2}\sigma} \right) - erf \left( \frac{x_{j} - \mu}{\sqrt{2}\sigma} \right) \right] \right\}.
\end{aligned}
\end{equation}

@@

%\newpage

\bibliography{digitisation}

\end{document}

>>>>

% mnras_template.tex
%
% LaTeX template for creating an MNRAS paper
%
% v3.0 released 14 May 2015
% (version numbers match those of mnras.cls)
%
% Copyright (C) Royal Astronomical Society 2015
% Authors:
% Keith T. Smith (Royal Astronomical Society)

% Change log
%
% v3.0 May 2015
%    Renamed to match the new package name
%    Version number matches mnras.cls
%    A few minor tweaks to wording
% v1.0 September 2013
%    Beta testing only - never publicly released
%    First version: a simple (ish) template for creating an MNRAS paper

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Basic setup. Most papers should leave these options alone.
\documentclass[a4paper,fleqn,usenatbib]{mnras}

% MNRAS is set in Times font. If you don't have this installed (most LaTeX
% installations will be fine) or prefer the old Computer Modern fonts, comment
% out the following line
\usepackage{newtxtext,newtxmath}
% Depending on your LaTeX fonts installation, you might get better results with one of these:
%\usepackage{mathptmx}
%\usepackage{txfonts}

\usepackage{multirow}

% Use vector fonts, so it zooms properly in on-screen viewing software
% Don't change these lines unless you know what you are doing
\usepackage[T1]{fontenc}
\usepackage{ae,aecompl}

@@

\newcommand{\tbd}[1]{\textcolor{Red}{{\bf TBD}: #1}}
\newcommand{\gab}[1]{\textcolor{Orchid}{[{\bf GS}: #1]}}
\newcommand{\changed}[1]{\textcolor{Red}{#1}}
\newcommand{\removed}[1]{\textcolor{Red}{}}

%%%%% AUTHORS - PLACE YOUR OWN PACKAGES HERE %%%%%

% Only include extra packages if you really need them. Common packages are:
\usepackage{graphicx}	% Including figure files
\usepackage{amsmath}	% Advanced maths commands
\usepackage{amssymb}	% Extra maths symbols

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%% AUTHORS - PLACE YOUR OWN COMMANDS HERE %%%%%

% Please keep new commands to a minimum, and use \newcommand not \def to avoid
% overwriting existing commands. Example:
%\newcommand{\pcm}{\,cm$^{-2}$}	% per cm-squared

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%% TITLE PAGE %%%%%%%%%%%%%%%%%%%

% Title of the paper, and the short title which is used in the headers.
% Keep the title short and informative.
\title[Short title, max. 45 characters]{Extreme Digitisation For Ground-Based Cosmic Microwave Background Experiments}

% The list of authors, and the short list which is used in the headers.
% If you need two or more lines of authors, add an extra line using \newauthor
\author[L. Balkenhol and C. L. Reichardt.]{
L. Balkenhol,$^{1}$\thanks{E-mail: lbalkenhol@student.unimelb.edu.au}
C. L. Reichardt,$^{1}$
\\
% List of institutions
$^{1}$School of Physics, University of Melbourne, Parkville, VIC 3010, Australia\\
}

% These dates will be filled out by the publisher
\date{Accepted XXX. Received YYY; in original form ZZZ}

% Enter the current year, for the copyright statements etc.
\pubyear{2019}

% Don't change these lines
\begin{document}
\label{firstpage}
\pagerange{\pageref{firstpage}--\pageref{lastpage}}
\maketitle

% Abstract of the paper
\begin{abstract}
The large size of the time ordered data of cosmic microwave background experiments presents challenges for mission planning and data analysis. These issues are particularly significant for Antarctica- and space-based experiments, which depend on satellite links to transmit data. We explore the viability of reducing the time ordered data to few bit numbers to address these challenges. Unlike lossless compression, few bit digitisation introduces additional noise into the data. We present a set of one, two, and three bit digitisation schemes and measure the increase in noise in the cosmic microwave background temperature and polarisation power spectra. The digitisation noise is independent of angular scale and is well-described as a constant percentage of the original detector noise. Three bit digitisation increases the map noise level by $< 2\%$, while reducing the data volume by a factor of ten relative to 32-bit floats. Extreme digitisation is a promising strategy for upcoming experiments.
\end{abstract}

% Select between one and six entries from the list of approved keywords.
% Don't make up new ones.
\begin{keywords}
cosmic background radiation -- polarization -- techniques: miscellaneous
\end{keywords}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%% BODY OF PAPER %%%%%%%%%%%%%%%%%%

\section{Introduction}
\label{sec:intro}

@@

As seen in Figure \ref{fig:psrecover} (appendix) all spectra of extremely digitised data recover the input to a satisfying degree. Differences between few bit and the input power spectra are due to observing a single CMB realisation and residual boundary effects. Non-linearity introduced by the digitisation process disappears in the final map due to the large number of TOD points that contribute to each map pixel. Thus the digitisation effects are not an issue for power spectrum estimation.

@@

Current-generation ground-based CMB experiments employ a combination of lossy, e.g. downsampling, and lossless compression, e.g. FLAC, bzip, LZMA, to manage their transmission bottlenecks. We test whether the data volume of timestreams that have undergone few-bit digitisation can be reduced further by applying the most popular lossless compression algorithm: FLAC.

We obtain 2 minutes of SPT-3G TOD and apply the digitisation schemes in equations \ref{eq:1bit}-\ref{eq:3bit} before compressing the data further using FLAC. We compare the file size of FLAC compressed few-bit timestreams to the typical data reduction rate achieved for 24-bit SPT-3G data.

We find that using FLAC in conjunction with extreme digitisation does not yield additional data volume reduction. For 3-bit digitisation, both the FLAC and raw timestream have $\sim$3 bits per data point. In fact, due to the overhead required by FLAC compressed files perform marginally worse than uncompressed few-bit streams. This is unsurprising, given that we have already maximally compressed the data using extreme digitisation. For comparison, FLAC reduces the full 24-bit  SPT-3G data to $6-8$ bits of entropy per number. For this real-world case, three-bit digitisation reduces the data volume by a factor of $2-3$ more than lossless compression.

\section{Conclusions}
\label{sec:conclusions}

@@

We present a set of one, two, and three bit digitisation schemes. For white and $1/f$ detector noise alike, we find that optimal three bit digitisation adds as little as $<2\%$ to the map noise level. This is true for temperature and polarisation observations. No change in the results is observed for maps with different numbers of hits per pixel and no angular scale dependence was observed in the added noise. Applying lossless compression techniques like FLAC to few-bit digitised data does not yield further reductions in data volume.

@@

\section*{Acknowledgements}

We are indebted to Nathan Whitehorn for the details about SPT data compression provided by him.
We are grateful for insightful discussions about the prospects of extreme digitisation for CMB experiments with Andrew Melatos and Patrick Clearwater.
We thank the referee as well as Nikhel Gupta, Srinivasan Raghunathan, Federico Bianchini, Andrew Melatos, and Patrick Clearwater for valuable feedback on the manuscript.
We acknowledge support from an Australian Research Council Future Fellowship (FT150100074), and also from the University of Melbourne. 
This research used resources of the National Energy Research Scientific Computing Center, which is supported by the Office of Science of the U.S. Department of Energy under Contract No. DE-AC02-05CH11231. 
We acknowledge the use of the Legacy Archive for Microwave Background Data Analysis (LAMBDA). Support for LAMBDA is provided by the NASA Office of Space Science.
This research made use of the NumPy \citep{numpy}, SciPy \citep{scipy}, Matplotlib \citep{matplotlib}, and Astropy \citep{astropy} packages.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%% REFERENCES %%%%%%%%%%%%%%%%%%

% The best way to enter references is to use BibTeX:

\bibliographystyle{mnras}
\bibliography{digitisation} % if your bibtex file is called example.bib

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%% APPENDICES %%%%%%%%%%%%%%%%%%%%%

\onecolumn

\appendix

\section*{Appendix}

% result tables
%\subsection{Digitisation Noise Levels}
%\label{subsec:appendixnoisetables}

\begin{figure*}\centering
\vspace{-1.5cm}
\includegraphics[width=0.8\textwidth,clip]{psrecovery.pdf}
  \caption[Current ]{
  Input power spectra used to generate the CMB template map plus detector noise level (black line) and digitisation noise (grey line); power spectra using three- (violet marker, right), two- (orange marker, middle), and one-bit (green marker, left) TOD. Power spectra are rescaled to $D_\ell = \ell (\ell+1) C_\ell /2\pi$ and bined to $\Delta \ell = 100$. Main panels show the TT (top), EE (middle), and BB (bottom) power spectra; accompanying panels show the difference between the input plus noise curves and the power spectra recovered from few-bit TOD. Spectra constructed from few-bit TOD recover the input well. This plot shows results for a simulation with $1,024,000$ hits per map pixel using Gaussian white detector noise.}
\label{fig:psrecover}

@@

\def\arraystretch{1.3}
\begin{table*}
\begin{center}
\caption{ \label{tab:extranoisewhite} Fractional Noise Increase Due To Digitisation - Gaussian White Detector Noise}
\small
\begin{tabular}{c c c c c}
Hits Per Pixel & Channel & 1 Bit & 2 Bit & 3 Bit \\
\hline
\hline
\multirow{3}{*}{800}  & TT  & $ 0.252 \pm 0.0091 $  & $ 0.0639 \pm 0.0034 $  & $ 0.0173 \pm 0.0021 $ \\
& EE  & $ 0.253 \pm 0.012 $  & $ 0.0640 \pm 0.0047 $  & $ 0.0175 \pm 0.0041 $ \\
& BB  & $ 0.254 \pm 0.011 $  & $ 0.0652 \pm 0.0054 $  & $ 0.0180 \pm 0.0039 $ \\
\hline
\multirow{3}{*}{8,000}  & TT  & $ 0.2521 \pm 0.0076 $  & $ 0.0641 \pm 0.0034 $  & $ 0.0174 \pm 0.0017 $ \\
& EE  & $ 0.2519 \pm 0.0074 $  & $ 0.0636 \pm 0.0040 $  & $ 0.0173 \pm 0.0022 $ \\
& BB  & $ 0.252 \pm 0.018 $  & $ 0.0634 \pm 0.0089 $  & $ 0.0176 \pm 0.0036 $ \\
\hline
\multirow{3}{*}{80,000}  & TT  & $ 0.2526 \pm 0.0080 $  & $ 0.0645 \pm 0.0040 $  & $ 0.0181 \pm 0.0016 $ \\
& EE  & $ 0.2506 \pm 0.0093 $  & $ 0.0641 \pm 0.0042 $  & $ 0.0177 \pm 0.0019 $ \\
& BB  & $ 0.250 \pm 0.012 $  & $ 0.0636 \pm 0.0065 $  & $ 0.0179 \pm 0.0032 $ \\
\hline
\multirow{3}{*}{1,024,000}  & TT  & $ 0.2501 \pm 0.0090 $  & $ 0.0633 \pm 0.0031 $  & $ 0.0173 \pm 0.0016 $ \\
& EE  & $ 0.2518 \pm 0.0080 $  & $ 0.0646 \pm 0.0033 $  & $ 0.0179 \pm 0.0018 $ \\
& BB  & $ 0.2520 \pm 0.0082 $  & $ 0.0639 \pm 0.0037 $  & $ 0.0173 \pm 0.0021 $ \\
\hline
\multirow{3}{*}{10,240,000}  & TT  & $ 0.2480 \pm 0.0078 $  & $ 0.0630 \pm 0.0038 $  & $ 0.0175 \pm 0.0018 $ \\
& EE  & $ 0.2526 \pm 0.0077 $  & $ 0.0647 \pm 0.0033 $  & $ 0.0183 \pm 0.0018 $ \\
& BB  & $ 0.252 \pm 0.010 $  & $ 0.0642 \pm 0.0042 $  & $ 0.0177 \pm 0.0020 $ \\
\hline
\multirow{3}{*}{102,400,000}  & TT  & $ 0.2465 \pm 0.0049 $  & $ 0.0629 \pm 0.0019 $  & $ 0.0170 \pm 0.0010 $ \\
& EE  & $ 0.2511 \pm 0.0078 $  & $ 0.0644 \pm 0.0035 $  & $ 0.0175 \pm 0.0018 $ \\
& BB  & $ 0.2523 \pm 0.0086 $  & $ 0.0640 \pm 0.0031 $  & $ 0.0179 \pm 0.0017 $ \\
\hline
\multirow{3}{*}{Average}  & TT  & $ 0.2502 \pm 0.0017 $  & $ 0.0636 \pm 0.0017 $  & $ 0.0175 \pm 0.0017 $ \\
 & EE  & $ 0.2518 \pm 0.0023 $  & $ 0.0643 \pm 0.0023 $  & $ 0.0177 \pm 0.0023 $ \\
 & BB  & $ 0.2519 \pm 0.0027 $  & $ 0.0640 \pm 0.0027 $  & $ 0.0177 \pm 0.0027 $ \\
\end{tabular}
\vspace{1mm}
\description{ 
Fractional addition to the map noise level, $\Delta \sigma / \sigma$, due to one-, two-, and three-bit digitisation for white detector noise with $1\sigma$ confidence intervals. Note that there is no appreciable variation between temperature and polarisation observations, or between different depths of observation. It is striking that three-bit digitisation only leads to a percent-level increase in the map noise level.}
 \normalsize
\end{center}
\end{table*}

\def\arraystretch{1.3}
\begin{table*}
\begin{center}
\caption{\label{tab:extranoiseoof} Fractional Noise Increase Due To Digitisation - $1/f$ Detector Noise}
\small
\begin{tabular}{c c c c c}
Hits Per Pixel & Channel & 1 Bit & 2 Bit & 3 Bit \\
\hline
\hline
\multirow{3}{*}{8,000}  & TT  & $ 0.231 \pm 0.017 $  & $ 0.0584 \pm 0.0052 $  & $ 0.0159 \pm 0.0022 $ \\
& EE  & $ 0.250 \pm 0.012 $  & $ 0.0641 \pm 0.0043 $  & $ 0.0177 \pm 0.0019 $ \\
& BB  & $ 0.245 \pm 0.021 $  & $ 0.0623 \pm 0.0067 $  & $ 0.0169 \pm 0.0032 $ \\
\hline
\multirow{3}{*}{102,400}  & TT  & $ 0.235 \pm 0.014 $  & $ 0.0597 \pm 0.0042 $  & $ 0.0166 \pm 0.0019 $ \\
& EE  & $ 0.2503 \pm 0.0091 $  & $ 0.0636 \pm 0.0042 $  & $ 0.0174 \pm 0.0016 $ \\
& BB  & $ 0.249 \pm 0.024 $  & $ 0.0641 \pm 0.0069 $  & $ 0.0178 \pm 0.0025 $ \\
\hline
\multirow{3}{*}{1,024,000}  & TT  & $ 0.238 \pm 0.013 $  & $ 0.0601 \pm 0.0046 $  & $ 0.0168 \pm 0.0019 $ \\
& EE  & $ 0.2502 \pm 0.0078 $  & $ 0.0630 \pm 0.0039 $  & $ 0.0173 \pm 0.0019 $ \\
& BB  & $ 0.2499 \pm 0.0095 $  & $ 0.0636 \pm 0.0049 $  & $ 0.0177 \pm 0.0019 $ \\
\hline
\multirow{3}{*}{Average}  & TT  & $ 0.2346 \pm 0.0020 $  & $ 0.0594 \pm 0.0020 $  & $ 0.0164 \pm 0.0020 $ \\
 & EE  & $ 0.2503 \pm 0.0018 $  & $ 0.0636 \pm 0.0018 $  & $ 0.0175 \pm 0.0018 $ \\
 & BB  & $ 0.2480 \pm 0.0025 $  & $ 0.0633 \pm 0.0025 $  & $ 0.0175 \pm 0.0025 $ \\
\end{tabular}
\vspace{1mm}
\description{ 
Fractional addition to the map noise level, $\Delta \sigma / \sigma$, due to one-, two-, and three-bit digitisation for $1/f$ detector noise with $1\sigma$ confidence intervals. As in the white noise case, we observe no significant trend between temperature and polarisation or for different depths of observation. We note that, even assuming a more realistic noise profile, three-bit digitisation only leads to a percent-level increase in the map noise level.}
 \normalsize
\end{center}
\end{table*}

@@

\begin{equation}
\langle D(\mu + \xi_1) D(\mu + \xi_2) \rangle = \left\{ \sum_{i=0}^N  \frac{y_i}{2} \left[ \mathrm{erf} \left( \frac{x_{i+1} - \mu}{\sqrt{2}\sigma} \right) - \mathrm{erf} \left( \frac{x_{i} - \mu}{\sqrt{2}\sigma} \right) \right] \right\} \times \left\{  \sum_{j=0}^N \frac{y_j}{2} \left[ \mathrm{erf} \left( \frac{x_{j+1} - \mu}{\sqrt{2}\sigma} \right) - \mathrm{erf} \left( \frac{x_{j} - \mu}{\sqrt{2}\sigma} \right) \right] \right\}.
\end{equation}

@@

% Don't change these lines
\bsp	% typesetting comment
\label{lastpage}
\end{document}

% End of mnras_template.tex
